---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preparation 

Check n in each condition:

```{r}
table(tdata$structure, tdata$test_queries)
```


Factorize:

```{r}
tdata$structure <- factor(tdata$structure, levels = c("irreversible", "reversible"), labels = c("irreversible", "reversible"))
tdata$test_queries <- factor(tdata$test_queries, levels = c("original", "novel"), labels = c("original", "novel"))
```


Make a subset containing only the columns relevant for analyses and turn into long format:

```{r}
# Subset: 
tdata_sub <- subset(tdata, select = c("subj_code", "condition", "structure", "test_queries", "first_caused_rating", "second_caused_rating",
                                      "first_maintained_rating", "second_maintained_rating"))

tdata_original <- subset(tdata_sub, test_queries == "original", select = c(1:6))
tdata_novel <- subset(tdata_sub, test_queries == "novel")

# into long format:
library(tidyr)
tdata_long_original <- gather(tdata_original, entity, rating, 5:6)
tdata_long_novel <- gather(tdata_novel, entity, rating, 5:8)

# factorize entity 
tdata_long_original$entity <- factor(tdata_long_original$entity, levels = c("first_caused_rating", "second_caused_rating"), 
                            labels = c("first \ncaused", "second \ncaused"))



tdata_long_novel$entity <- factor(tdata_long_novel$entity, levels = c("first_caused_rating", "second_caused_rating", 
                                                          "first_maintained_rating", "second_maintained_rating"), 
                            labels = c("first \ncaused", "second \ncaused","first \nmaintained", "second \nmaintained"))


tdata_long_novel$rating <- as.numeric(tdata_long_novel$rating)

```


# Replication part of study 

## Graphs and summaries

Replication part of the study 

```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


tdata_sub <- tdata_long_original


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))



g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ structure, labeller = label_both)+
  scale_y_continuous(limits = c(-0.3, 8.3), breaks=seq(0, 8, 1), expand = c(0,0)) +
  geom_violinhalf(aes(y = rating, group = entity, fill = structure), color = NA, 
                position=position_dodge(1), alpha = 0.4)+
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, color = structure), fun.y=mean, geom="line", 
               shape = 22, size = 1.5, alpha = .7)+
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 2, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 8.2, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -2)

g

#ggsave("results_means_mainDV_original.svg",width=8,height=5)
#ggsave("results_means_mainDV_original.pdf",width=8,height=5)

```




```{r}
library(ggridges)
g2 <- ggplot(tdata_sub, aes(x = rating, y = entity, fill = entity)) +
  facet_grid( ~ structure,labeller = label_both)+
  scale_x_continuous(breaks = seq(0, 10, 1))+
  geom_density_ridges(alpha = 0.5)+
   #stat_summary(aes(x = rating_rec), fun.x=mean, geom="point", 
  #             color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_viridis_c(name = "Explanation \nRating", option = "C", breaks=c(-5,0,5), labels=c("narrow scope", "no preference", "broad scope"))+
  labs(x = "Agreement rating", y = "Causal Statement") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  theme(axis.text.y = element_text(size = 14, angle = 0))

g2

#ggsave("results_dist_original.svg",width=8,height=5)
#ggsave("results_dist_original.pdf",width=8,height=5)
```


## Mixed ANOVA 

```{r}
library(afex)
library(emmeans)

a1 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_long_original, 
              anova_table = list(es = "pes"))
a1

```

No significant interaction between structure and entity (causal statement), but a clear effect of entity (causal statement). This replicates the original study.


## Contrasts 

```{r}

library(lsmeans)
# means

ls2 <- lsmeans(a1, c("structure", "entity")) 
ls2

contrasts <- emmeans(a1, ~ entity|structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```
First cause receives higher ratings in both reversibility conditions. 


# Novel part of the study 

```{r}

tdata_sub <- tdata_long_novel


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))



g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ structure, labeller = label_both)+
  scale_y_continuous(limits = c(-0.3, 8.3), breaks=seq(0, 8, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, color = structure), fun.y=mean, geom="line", 
               shape = 22, size = 1.5, alpha = .7)+
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g

#ggsave("results_means_mainDV_novel.svg",width=12,height=5)
#ggsave("results_means_mainDV_novel.pdf",width=12,height=5)

```



```{r}
library(ggridges)
g2 <- ggplot(tdata_sub, aes(x = rating, y = entity, fill = structure)) +
  facet_grid( ~ structure,labeller = label_both)+
  scale_x_continuous(breaks = seq(0, 10, 1))+
  geom_density_ridges(alpha = 0.5)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  labs(x = "Agreement Rating", y = "Causal Statement") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  theme(axis.text.y = element_text(size = 14, angle = 0))+
  annotate("text", y = 0.7, x = 0, label = c("completely disagree"), angle = 0)+
  annotate("text", y = 0.7, x = 9, label = c("completely agree"), angle = 0)

g2

#ggsave("results_dist_novel.svg",width=12,height=5)
#ggsave("results_dist_novel.pdf",width=12,height=5)
```


## Mixed ANOVA 

```{r}
library(afex)
library(emmeans)

a2 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_long_novel, anova_table = list(es = "pes"))
a2

```
Predicted interaction effect was obtained, but also visualize the main effects: 

Plot main effect of structure: 

```{r}

tdata_sub <- tdata_long_novel


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = structure, y = valueJitter)) +
  guides(fill=FALSE)+
  scale_y_continuous(limits = c(-0.3, 8.3), breaks=seq(0, 8, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g

```

Plot main effect of entity (causal statement):

```{r}

tdata_sub <- tdata_long_novel


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  scale_y_continuous(limits = c(-0.3, 8.3), breaks=seq(0, 8, 1), expand = c(0,0)) +
  geom_jitter(aes(color = entity), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = entity), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g
```


## Contrasts 

```{r}

library(lsmeans)
# means

ls2 <- lsmeans(a2, c("structure", "entity")) 
ls2

contrasts <- emmeans(a2, ~ entity|structure)
s <- pairs(contrasts, adjust = "none")


s
confint(s, level = 0.95)
```

The planned contrasts from this list are "first caused - second caused" in each reversibility condition. They are both significant. 


The other planned contrast is the one comparing the maintainer ratings between the two reversibility conditions. The following code will produce a long list of comparisons, from which the relevant one will be "second..maintained irreversible - second..maintained reversible"


```{r}
contrasts <- emmeans(a2, ~ entity*structure)
s <- pairs(contrasts, adjust = "none")


s
confint(s, level = 0.95)
```

The relevant contrast "second..maintained irreversible - second..maintained reversible" is significant


