---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(ez)
library(data.table)
library(tidyverse) 

library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```


```{r}
tdata <- read_csv("exp_data.csv")
```


```{r}
# Exclude subjects who failed the control questions 
tdata <- subset(tdata, intro_check == "2: A little squid swam from the right to the left.")
tdata <- subset(tdata, timing_check_correct == "correct")
tdata <- subset(tdata, learning_check_correct == "correct")
```


# Subject demographics

```{r}
# demographics 

min(tdata$age)
max(tdata$age)
mean(tdata$age)
sd(tdata$age)

# 1 = male, 2 = female, 3 = other
table(tdata$gender)
```

1 = male, 2 = female, 3 = non-binary, 4 = prefer not to say


# Data preparations 

Check n in each condition:

```{r}
table(tdata$structure, tdata$first_object)
```


Factorize:

```{r}
tdata$structure <- factor(tdata$structure, levels = c("irreversible", "reversible"), labels = c("irreversible", "reversible"))
```

Relabel rating columns for red and yellow entity according to "first object" (so that they are called first_obj_rating and second_obj_rating):

```{r}
left_first <- subset(tdata, first_object == "left")
colnames(left_first)[10] ="first_obj_caused_rating"
colnames(left_first)[11] ="second_obj_caused_rating"
colnames(left_first)[12] ="first_obj_maintained_rating"
colnames(left_first)[13] ="second_obj_maintained_rating"

right_first <- subset(tdata, first_object == "right")
colnames(right_first)[11] ="first_obj_caused_rating"
colnames(right_first)[10] ="second_obj_caused_rating"
colnames(right_first)[13] ="first_obj_maintained_rating"
colnames(right_first)[12] ="second_obj_maintained_rating"

right_first <- right_first[c("subj_code", "condition", "desktop_conf", "attent_conf", "intro_check", 
                             "structure", "first_object", "learn_check_left", "learn_check_right",
                             "first_obj_caused_rating", "second_obj_caused_rating", "first_obj_maintained_rating", "second_obj_maintained_rating",
                             "timing_check", "recorded_at", 
                             "age", "gender", "tech_issues", "timing_check_correct", "learning_check_left_correct", 
                             "learning_check_right_correct", "learning_check_correct")]


names(left_first) %in% names(right_first)

tdata <- rbind(left_first, right_first)
```

Make a subset containing only the columns relevant for analyses and turn into long format:

```{r}
# Subset: 
tdata_sub <- subset(tdata, select = c("subj_code","structure","first_obj_caused_rating", "second_obj_caused_rating",
                                      "first_obj_maintained_rating", "second_obj_maintained_rating"))

# into long format:
library(tidyr)
tdata_long <- gather(tdata_sub, entity, rating, 3:6)


# factorize entity 
tdata_long$entity <- factor(tdata_long$entity, levels = c("first_obj_caused_rating", "second_obj_caused_rating", 
                                                          "first_obj_maintained_rating", "second_obj_maintained_rating"), 
                            labels = c("first \ncaused", "second \ncaused","first \nmaintained", "second \nmaintained"))
```



# Graphs and summaries


```{r}
myTheme <- theme(plot.title = element_text(face="bold", size = 22),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text.x = element_text(size = 14, angle = 0), 
        axis.text.y = element_text(size = 16, angle = 0),
        legend.text = element_text(size = 18),
        legend.title = element_text(face = "bold", size = 18),
        strip.text.x = element_text(size = 18),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))


tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  facet_grid( ~ structure,labeller = label_both)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, color = structure), fun.y=mean, geom="line", 
               shape = 22, size = 1.5, alpha = .7)+
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)+
  annotate("text", x = 0.5, y = 2, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 8.2, label = c("completely agree"), angle = 90)

g

#ggsave("results_means_mainDV.svg",width=12,height=5)
#ggsave("results_means_mainDV.pdf",width=12,height=5)

```




```{r}
library(ggridges)
g2 <- ggplot(tdata_long, aes(x = rating, y = entity, fill = structure)) +
  facet_grid( ~ structure,labeller = label_both)+
  scale_x_continuous(breaks = seq(0, 10, 1))+
  geom_density_ridges(alpha = 0.5)+
   #stat_summary(aes(x = rating_rec), fun.x=mean, geom="point", 
  #             color = "black", shape = 22, size = 2, group=1, alpha = 1)+
  scale_fill_manual(values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  #scale_fill_viridis_c(name = "Explanation \nRating", option = "C", breaks=c(-5,0,5), labels=c("narrow scope", "no preference", "broad scope"))+
  labs(x = "Agreement Rating", y = "Causal Statement") +
  scale_y_discrete(limits=rev)+
  myTheme+
  theme_light(base_family = "Poppins", base_size = 20)+
  theme(panel.grid = element_blank(), axis.text = element_text(colour ="black"))+
  theme(legend.position="none",
        legend.title=element_blank(),legend.key.width = unit(1.95, 'cm'))+
  theme(axis.text.y = element_text(size = 14, angle = 0))+
  annotate("text", y = 0.7, x = 0, label = c("completely disagree"), angle = 0)+
  annotate("text", y = 0.7, x = 10, label = c("completely agree"), angle = 0)

g2

#ggsave("results_dist.svg",width=12,height=5)
#ggsave("results_dist.pdf",width=12,height=5)
```


## Mixed ANOVA 

```{r}
library(afex)
library(emmeans)

a1 <- aov_car(rating ~ structure*entity + Error(subj_code/entity), tdata_long, anova_table = list(es = "pes"))
a1

```
Predicted interaction effect was significant-


Visualize also the obtained main effect of "structure" (reversibility)

```{r}
tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))


g <- ggplot(tdata_sub, aes(x = structure, y = valueJitter)) +
  guides(fill=FALSE)+

  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = structure), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = structure), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)+
  annotate("text", x = 0.5, y = 2, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 8.2, label = c("completely agree"), angle = 90)

g

```

Plot main effect of entity (causal statement):

```{r}

tdata_sub <- tdata_long


library(see)
## first, turn sID into a factor
tdata_sub$subj_code <- factor(tdata_sub$subj_code)

pd <- position_dodge(width = 0.3)

tdata_sub$valueJitter <- jitter(tdata_sub$rating, factor = 0.01, amount = 0.004)

theme_set(theme_light(base_size = 20, base_family = "Poppins"))

g <- ggplot(tdata_sub, aes(x = entity, y = valueJitter)) +
  guides(fill=FALSE)+
  scale_y_continuous(limits = c(-0.3, 10.3), breaks=seq(0, 10, 1), expand = c(0,0)) +
  geom_jitter(aes(color = entity), alpha = 0.5, width = 0.15, height = 0.2) +
  stat_summary(aes(y = rating, group=1), fun.data = mean_cl_boot, 
               geom = "errorbar", width = 0, size = 1) +
  stat_summary(aes(y = rating, group=1, fill = entity), fun.y=mean, geom="point", 
               color = "black", shape = 22, size = 3, group=1, alpha = 1)+
  stat_summary(aes(y = rating,group=1), fun.y=median, geom="point", color = "black", shape = 3, size = 4, 
               group=1, alpha = 1, position = position_dodge(width = 0.5))+
  labs(x = "Causal Statement", y = "Agreement Rating") +
  scale_color_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  scale_fill_manual(name = "Strength",values=c("#66c2a5", "#e78ac3", "#8da0cb", "#a6d854"))+
  theme(legend.position = "none")+
  annotate("text", x = 0.5, y = 1.7, label = c("completely disagree"), angle = 90)+
  annotate("text", x = 0.5, y = 6.7, label = c("completely agree"), angle = 90)+
  myTheme+
  theme(panel.grid.major = element_line(color = "lightgrey",
                                        size = 0.5,
                                        linetype = 'dotted'))+
  stat_summary(aes(label=round(after_stat(y),2)), fun.y=mean, geom="text", size=5,
             vjust = -6)

g


```


## Contrasts 



```{r}

library(lsmeans)
# means

ls2 <- lsmeans(a1, c("structure", "entity")) 
ls2

contrasts <- emmeans(a1, ~ entity|structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The planned contrasts from the list above are " first..caused - second..caused" in both "irreversible" and "reversible". Both are significant.

Get effect standardized effect sizes for the contrasts

```{r}
# using the d.dep.t.avg function from the MOTE package (see https://matthewbjane.quarto.pub/guide-to-effect-sizes-and-confidence-intervals/Standardized-Mean-Differences.html#sec-repeated-measures-drm)

library(MOTE)

# from the contrast outputs above, get the SDs for pre and post. 

# irreversible condition
# sds required for function can be obtained by multiplying the SE with the square root of the sample size

#structure    entity             lsmean    SE df lower.CL upper.CL
# irreversible first..caused        9.64 0.178 98    9.286     9.99
# irreversible second..caused       0.92 0.408 98    0.110     1.73



stats <- d.dep.t.avg(
  m1 = 9.64,
  m2 = 0.92,
  sd1 = 0.178 * sqrt(100),
  sd2 = 0.408 * sqrt(100),
  n = 100,
  a = 0.05
)

stats$estimate
```

```{r}
# reversible condition 
# sds required for function can be obtained by multiplying the SE with the square root of the sample size

#structure    entity             lsmean    SE df lower.CL upper.CL
# reversible   first..caused        9.40 0.178 98    9.046     9.75
# reversible   second..caused       2.74 0.408 98    1.930     3.55

stats <- d.dep.t.avg(
  m1 = 9.40,
  m2 = 2.74,
  sd1 = 0.178 * sqrt(100),
  sd2 = 0.408 * sqrt(100),
  n = 100,
  a = 0.05
)

stats$estimate
```



The final planned contrast compares the maintainer ratings for the second cause in "irreversible" and "reversible". It must be looked up from the long list of contrasts below: 

```{r}
contrasts <- emmeans(a1, ~ entity*structure)
s <- pairs(contrasts, adjust = "none") 


s
confint(s, level = 0.95)
```

The relevant contrast is: "second..maintained irreversible - second..maintained reversible", which is significant. 

Get effect size: 

```{r}
# reversible condition 
# sds required for function can be obtained by multiplying the SE with the square root of the sample size

#structure    entity             lsmean    SE df lower.CL upper.CL
#irreversible second..maintained   3.32 0.409 98    2.508     4.13
#reversible   second..maintained   9.36 0.409 98    8.548    10.17

stats <- d.dep.t.avg(
  m1 = 9.36,
  m2 = 3.32,
  sd1 = 0.409 * sqrt(100),
  sd2 = 0.409 * sqrt(100),
  n = 100,
  a = 0.05
)

stats$estimate
```

